### 2021-12-09 paper/paper.tex
[paper]: add latex shell and makefile

\documentclass{article}

\usepackage[letterpaper,top=2cm,bottom=2cm,left=2cm,right=2cm,marginparwidth=1.75cm]{geometry}

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}

\pdfpagewidth=8.5in
\pdfpageheight=11in

\title{Compiler and Runtime Support for Functions-as-a-Service (FaaS)}
\date{}
\author{Souradip Ghosh, 15-849, Fall 2021}

\begin{document}
\maketitle

\section{Abstract}

\section{Introduction/Background}

\section{Design}

\section{Implementation}

\section{Evaluation}

\section{Future Work}

\section{Conclusion}

\bibliographystyle{plain}
\bibliography{sample}

\end{document}

### 2021-12-09 passes/scripts/pass_build.sh

@@ -10,10 +10,11 @@ mkdir -p ${TC}/catpass/src ;
 
 
 # Copy the sources into the toolchain for build
-cp -a ${PASS_DIR}/${DRIVER_FILE} ${TC}/catpass/ ;
-cp -a ${PASS_DIR}/include/ ${TC}/catpass/ ;
-cp -a ${PASS_DIR}/src/ ${TC}/catpass/ ;
-cp -a ${PASS_DIR}/CMakeLists.txt ${TC}/catpass/ ;
+cp -a ${PASS_DIR}/* ${TC}/catpass/ ;
+# cp -a ${PASS_DIR}/${DRIVER_FILE} ${TC}/catpass/ ;
+# cp -a ${PASS_DIR}/include/ ${TC}/catpass/ ;
+# cp -a ${PASS_DIR}/src/ ${TC}/catpass/ ;
+# cp -a ${PASS_DIR}/CMakeLists.txt ${TC}/catpass/ ;
 
 
 # Build from the toolchain

### 2021-12-09 paper/paper.tex
[paper]: add abstract

@@
 
+\def\crcd{compiler-runtime co-design}
+
 \title{Compiler and Runtime Support for Functions-as-a-Service (FaaS)}
@@
 \section{Abstract}
+Functions-as-a-Service (FaaS) are rapidly growing in popularity and 
+complexity among cloud service provides as users look to take 
+advantage of cloud resources on-demand and focus on computation
+at a finer grain. FaaS instances are characteristically short-lived
+and often dominated by communication and cold-start overheads compared
+to typical monolithic applications. At the same time, FaaS instances
+are requiring more resources, especially memory, to operate. Although
+many works have approached these issues with scheduling optimizations,
+hardware accelerators, and other techniques, very few have examined
+\textit{compiler} support and \textit{\crcd} for FaaS. This paper presents 
+a \crcd\ consisting of static analyses, profiling, and a custom memory 
+allocator controlled by the compiler and runtime to better understand 
+and optimize \textit{dynamic memory usage} for small FaaS workloads.
 

### 2021-12-09 paper/paper.tex
[paper]: add design overview, add system figure

@@
 \usepackage{graphicx}
+\usepackage{titlesec}
+\usepackage[parfill]{parskip}
 \usepackage[colorlinks=true, allcolors=blue]{hyperref}
@@
 \pdfpageheight=11in
+
+\def\tricolfig#1{\includegraphics[width=2.2in]{#1}}
+\def\triplecolfig#1{\includegraphics[width=0.32\textwidth]{#1}}
+\def\smallcolfig#1{\includegraphics[width=2.0in]{#1}}
+\def\smallercolfig#1{\includegraphics[width=0.75\columnwidth]{#1}}
+\def\mediumcolfig#1{\includegraphics[width=0.9\columnwidth]{#1}}
+\def\colfig#1{\includegraphics[width=\columnwidth]{#1}}
+\def\pagefig#1{\includegraphics[width=0.75\textwidth]{#1}}
+\def\suspcolfig#1{\includegraphics[width=2.0in]{#1}}
+\def\perfcolfig#1{\smallercolfig{#1}}
 
@@
 \section{Design}
+The design of the system consists of three major components : 1) a \textit{profiler},
+2) \textit{static analyses} in the compiler, and 3) a \textit{custom memory allocator}.
+The first and third components consist of compiler instrumentation coupled with an 
+integrated runtime. These compoents are described in the following subsections
+in detail.
+
+\begin{figure}
+    \centerline{\pagefig{figs/sys.pdf}}
+    \caption{System overview and compilation pipeline. }  
+	\label{fig:sys}
+\end{figure}
+
+An overview of the system and compilation pipeline is show in Figure~\ref{fig:sys}. The 
+first pass over an example workload (\texttt{app.c}) consists of static analyses and 
+profiling. The second pass over \texttt{app.c} instruments the program with the custom
+allocator using the results from the static analyses (and optionally the profiler). The
+outputs from this compilation pipeline include an instrumented application (\texttt{app.exe})
+statistics about dynamic memory working set size.
+
+\subsection{Profiler}
+
+\subsection{Static Analyses}
+
+\subsection{Custom Allocator}
 

### 2021-12-09 paper/paper.tex
[paper]: add first cut for profiler subsection

@@
 \def\crcd{compiler-runtime co-design}
+\def\ALL{allocation}
+\def\DALL{deallocation}
+\def\ALLS{allocations}
+\def\DALLS{deallocations}
+\def\DWS{dynamic memory working set size}
 
@@
 The first and third components consist of compiler instrumentation coupled with an 
-integrated runtime. These compoents are described in the following subsections
-in detail.
+integrated runtime. All analyses and transformations operate on the \textit{intermediate 
+representation} (IR) of the compiler, where standard compiler optimizations are performed.
+These components are described in the following subsections in detail.
 
@@
 outputs from this compilation pipeline include an instrumented application (\texttt{app.exe})
-statistics about dynamic memory working set size.
+and statistics about \DWS\.
 
 \subsection{Profiler}
+The profiler has the goal of understanding information about dynamic memory \ALLS\ and 
+\DALLS\ during execution (for instance, the number of \ALLS\, \DWS\, common \ALL 
+sizes, etc.) as well as standard information about the "hotness" of functions and
+loops at runtime. Using this data, the profiler can inform the compiler during 
+instrumentation in the second pass over the application (see \ref{ca}). It can also 
+inform the user with a better estimate of memory usage for a future FaaS instance or 
+function invocation.
+
+The profiler tracks all \ALLS and \DALLS of some workload, $W$, at runtime. This is 
+accomplished by instrumenting all explicit dynamic memory \ALL\ and \DALL\ function 
+calls in the IR for $W$ with \textit{callback} functions to a profiler runtime.
+Two callbacks, \texttt{TrackAllocation(ptr, size)} and \texttt{TrackDeallocation(ptr)}, are 
+sufficient to achieve all tracking. \texttt{TrackAllocation} takes a pointer to an allocation (\texttt{ptr}) and 
+the size of the allocation (\texttt{size}) as inputs and records the pair in a table in the 
+profiler's runtime. \texttt{TrackDeallocation} simply takes a pointer to an allocation (\texttt{ptr}) as
+input and removes the entry from the profiler runtime's table.
+
+While $W$ runs, the profiler tracks each change \DWS\ over its entire execution. Using 
+this information, the profiler calculates maximum and average \DWS\ for $W$ and outputs
+the data to the user and to the compiler for future use.
 

### 2021-12-09 allocator/include/RuntimeConfigurations.hpp
[runtime][allocator]: add timing macros for analysis

@@ -15,3 +15,13 @@
 #define RUNTIME_ASSERT_ON 0
 #define RUNTIME_ASSERT if (RUNTIME_ASSERT_ON) assert
 
+#define TIME 1
+#define TIME_START(Start) \
+    uint64_t Start; \
+    if (TIME) Start = rdtsc();
+    
+#define TIME_END_AND_PRINT(Prefix, Start) \
+    if (TIME) { \
+        uint64_t End = rdtsc() - Start; \
+        std::cerr << Prefix << End << std::endl; \
+    }

### 2021-12-09 allocator/src/FaaSAllocator.cpp

@@ -26,8 +26,10 @@ void Init(uint64_t PoolSize)
      * 
      * Build the compiler directed pool and set it to Allocator::TheAllocator
      */
+    TIME_START(Start);
     Allocator::TheAllocator = new Allocator::FaaSAllocator();
     Allocator::TheAllocator->Init(PoolSize);
+    TIME_END_AND_PRINT("Init: ", Start);
     return;
 }
 
@@ -44,10 +46,12 @@ void AddAllocator(
      */
     RUNTIME_DEBUG << "AddAllocator: [BumpIDBlockSize : PoolSize]: ["
                   << BumpIDBlockSize << " : " << PoolSize << "]" << std::endl;
+    TIME_START(Start);
     Allocator::TheAllocator->AddBumpAllocator(
         BumpIDBlockSize,
         PoolSize
     );
+    TIME_END_AND_PRINT("AddAllocator: ", Start);
 
 
     return;
@@ -61,7 +65,9 @@ void *AllocateFromCompilerDirectedPool(uint64_t Offset)
      *
      * Return a pointer that is offset @Offset into Allocator::CompilerPartitionedPool
      */
+    TIME_START(Start);
     uint64_t Address = ((uint64_t) Allocator::CompilerPartitionedPool) + Offset;
+    TIME_END_AND_PRINT("AllocateFromCompilerDirectedPool: ", Start);
     RUNTIME_DEBUG << "AllocateFromCompilerDirectedPool: [Address : Offset]: ["
                   << Address << " : " << Offset << "]" << std::endl;
     return ((void *) Address);
@@ -75,7 +81,9 @@ void *AllocateWithRuntimeInit(uint64_t BumpIDBlockSize)
      * 
      * Wrapper to allocate memory w/ runtime init in Allocator::TheAllocator 
      */
+    TIME_START(Start);
     void *Allocation = Allocator::TheAllocator->AllocateFromBumpWithRuntimeInit(BumpIDBlockSize);
+    TIME_END_AND_PRINT("AllocateWithRuntimeInit: ", Start);
     RUNTIME_DEBUG << "AllocateWithRuntimeInit: " << BumpIDBlockSize << std::hex << ", " << Allocation << std::dec << std::endl;
     return Allocation;
 }
@@ -89,7 +97,9 @@ void *Allocate(uint64_t BumpIDBlockSize)
      * 
      * Wrapper to allocate memory in Allocator::TheAllocator 
      */
[…]

### 2021-12-09 paper/paper.tex
[paper]: add more to profiler subsection, add example profiler instrumentation fig

@@
 \usepackage{titlesec}
+\usepackage{lipsum}
+\usepackage{mwe}
 \usepackage[parfill]{parskip}
@@
 outputs from this compilation pipeline include an instrumented application (\texttt{app.exe})
-and statistics about \DWS\.
+and statistics about \DWS\ (total size for all dynamic \ALLS).
 
@@
 The profiler tracks all \ALLS and \DALLS of some workload, $W$, at runtime. This is 
-accomplished by instrumenting all explicit dynamic memory \ALL\ and \DALL\ function 
+accomplished by using the compiler to instrument all explicit dynamic memory \ALL\ and \DALL\ function 
 calls in the IR for $W$ with \textit{callback} functions to a profiler runtime.
@@
 profiler's runtime. \texttt{TrackDeallocation} simply takes a pointer to an allocation (\texttt{ptr}) as
-input and removes the entry from the profiler runtime's table.
+input and removes the entry from the profiler runtime's table. An example of this 
+instrumentation is shown in Figure~\ref{fig:pins}.
 
-While $W$ runs, the profiler tracks each change \DWS\ over its entire execution. Using 
-this information, the profiler calculates maximum and average \DWS\ for $W$ and outputs
-the data to the user and to the compiler for future use.
+\begin{figure}
+    \centering
+    \begin{minipage}{0.45\textwidth}
+        \centering
+        \includegraphics[width=0.9\textwidth]{figs/pins.png} 
+        \caption{Example compiler instrumentation for profiler callbacks in the 
+        context of simplified LLVM IR and \textit{libc} \texttt{malloc}
+        and \texttt{free}}  
+	    \label{fig:pins}
+    \end{minipage}\hfill
+\end{figure}
+
+While $W$ runs, the profiler tracks each change to the \DWS\ over its entire execution 
+whenever a callback is invoked. Using this information, the profiler calculates maximum 
+and average \DWS\ for $W$ and outputs the data to the user and to the compiler for future 
+use.
 

### 2021-12-09 paper/paper.tex
[paper]: add subsection on static analyses

@@
 \def\DWS{dynamic memory working set size}
+\def\SIZE{allocation size}
+\def\SIZES{allocation sizes}
 
@@
 The design of the system consists of three major components : 1) a \textit{profiler},
-2) \textit{static analyses} in the compiler, and 3) a \textit{custom memory allocator}.
-The first and third components consist of compiler instrumentation coupled with an 
-integrated runtime. All analyses and transformations operate on the \textit{intermediate 
-representation} (IR) of the compiler, where standard compiler optimizations are performed.
-These components are described in the following subsections in detail.
+2) \textit{static analyses} in the compiler, and 3) a \textit{custom memory allocator}
+that is \textit{designed and optimized with FaaS characteristics in mind}. The first and third components consist 
+of compiler instrumentation coupled with an integrated runtime. All analyses and 
+transformations operate on the \textit{intermediate representation} (IR) of the compiler, 
+where standard compiler optimizations are performed. These components are described in 
+the following subsections in detail.
 
@@
 profiling. The second pass over \texttt{app.c} instruments the program with the custom
-allocator using the results from the static analyses (and optionally the profiler). The
+allocator using the results from the static analyses (and optionally the profiler). This 
+custom allocator replaces standard dynamic \ALLS\ and \DALLS\ in \texttt{app.c}, and it is
+the \textit{crux} of optimization for dynamic memory usage in FaaS workloads. The
 outputs from this compilation pipeline include an instrumented application (\texttt{app.exe})
@@
 and average \DWS\ for $W$ and outputs the data to the user and to the compiler for future 
-use.
+use. All profiler information can be optionally used by the custom allocator component of
+the \crcd\ (see \ref{ca}).
 
 \subsection{Static Analyses}
+The system also supports static analyses of \ALLS\ and \DALLS\ in the IR for workload $W$. 
+In particular, the compiler analyzes 1) \ALL\ sizes for all \ALLS\, 2) loops to unroll 
+and functions to inline so that 1) is more feasible to analyze, and 3) uses of \ALL\ and 
+\DALLS . All analyses informs the compiler about analyzable static allocations.
 
-\subsection{Custom Allocator}
+The first analysis is especially important to identify \ALL\ and \DALL\ candidates that
+can be replaced with the custom allocator. At a high level, the "easier it is to deduce"
+the allocation \textit{size} for each \ALL\, the easier it is to replace the \ALL\ with one
+of the methods from the custom alloactor (see \ref{ca} for details).
+
+An "easy-to-deduce" or \textit{statically analyzable} \SIZE\ is one that is 
+simply a constant integer. All other \SIZES\ are \textit{analyzable with
+assistance}, where the size can apparently be deduced during runtime. These 
+categories are mutually exclusive. The compiler records all statically analyzable 
+\SIZES .
+
+Some \SIZES\ that are analyzable with assistance are worth recording, however. In
+particular, if an \SIZE\ is dependent on function arguments, the function (and
+its parent functions) can be continually inlined until the size is statically
+analyzable. Consequently, the compiler marks functions that are candidates for
+aggressive inlining (compared to a standard inliner in \texttt{-O3}). At an
+extreme, aggressive inlining can reach a point where it may be beneficial to
+inline the entry-point function of $W$. This is especially common in small 
+FaaS workloads where \ALLS\ towards the beginning of the workload havew \SIZES\
+dependent on the inputs. Obviously, inlining the entry-point function is not 
+possible, but we can borrow concepts from JIT compilation or binary translation
+to possibly embed the inputs into the binary itself. This process can occur
+right before the $W$ is invoked in the FaaS instance. A subset of this system's compiler
+can analyze and transform $W$ as necessary. This solution is theoretical and 
+not applied in this work.
+
+Additionally, \SIZES\ that exists in a loop and are dependent on loop-invariant
+values are also recorded. Here, the loop itself can be aggressively unrolled and
+the \ALLS\ can be merged or replaced with one of the methods from the custom 
+allocator. It is also important that the size is dependent on a loop-invariant 
+value, or else the compiler cannot affirmatively analyze multiples of the size or
+determine an appropriate unroll factor (unrolling too much without knowledge of
+the \SIZE\ may be sub-optimal).
+
+Finally, the compiler analyzes the uses of \ALLS , especially if they are used
+in explicit \texttt{store} operations in the IR. An \ALL\ pointer that is stored
+to an arbitrary memory location is an \textit{escape}, and it is very difficult
+to analyze its value, lifetime, and size once it leaves the scope of its parent
+function. Additionally, it is hard to manage \DALLS\ if pointer origins
+cannot be deduced, which can be common with escaped \ALL\ pointers. All escapes
+are recorded. All \ALLS\ and \DALLS\ that cannot be statically analyzed using
+memory alias analysis (i.e "must alias" conditions) or def-use chains are flagged.
+
+\subsection{Custom Allocator} \label{ca}
 

### 2021-12-09 paper/paper.tex
[paper]: add figures for static analysis and pool transformation, start allocator subsection, fix content

@@
 \DALLS\ during execution (for instance, the number of \ALLS\, \DWS\, common \ALL 
-sizes, etc.) as well as standard information about the "hotness" of functions and
-loops at runtime. Using this data, the profiler can inform the compiler during 
+sizes, etc.) as well as typical information about critical paths or the "hotness" of functions and
+loops at runtime that is often use to aid optimization [?]. Using this data, the profiler can inform the compiler during 
 instrumentation in the second pass over the application (see \ref{ca}). It can also 
@@
     \end{minipage}\hfill
+    \begin{minipage}{0.45\textwidth}
+        \centering
+        \includegraphics[width=0.77\textwidth]{figs/sizes.png} 
+        \caption{Example allocations that the static analysis can target in the 
+        context of simplified LLVM IR and \textit{libc} \texttt{malloc}}  
+	    \label{fig:sizes}
+    \end{minipage}\hfill
 \end{figure}
@@
 
-\subsection{Static Analyses}
+\subsection{Static Analyses} \label{sa}
 The system also supports static analyses of \ALLS\ and \DALLS\ in the IR for workload $W$. 
@@
 categories are mutually exclusive. The compiler records all statically analyzable 
-\SIZES .
+\SIZES . An example is shown in Line 2 of Figure~\ref{fig:sizes}.
 
@@
 analyzable. Consequently, the compiler marks functions that are candidates for
-aggressive inlining (compared to a standard inliner in \texttt{-O3}). At an
+aggressive inlining (compared to a standard inliner in \texttt{-O3}). An example of
+this is shown in Line 3 of Figure ~\ref{fig:sizes}. At an
 extreme, aggressive inlining can reach a point where it may be beneficial to
@@
 determine an appropriate unroll factor (unrolling too much without knowledge of
-the \SIZE\ may be sub-optimal).
+the \SIZE\ may be sub-optimal). An example of this is shown in Line 8 of Figure ~\ref{fig:sizes}
 
@@
 \subsection{Custom Allocator} \label{ca}
+The final component of the system is a custom memory allocator, managed by both
+the compiler and runtime and designed with consideration of FaaS characteristics.
+In particular, the allocator aims to reduce raw \ALL\ and \DALL\ time, consolidate 
+\ALLS\ , and utilize compiler analysis to optimize dynamic memory overheads.
+
+The allocator primarly takes the static analyses (see \ref{sa})
+as input to replace as many \ALLS\ and \DALLS\ with invocations to the allocator. 
+First, \ALLS\ with statically analyzable \SIZES\ are prime candidates to be 
+\textit{pre-allocated} because all the information needed to allocate is known at
+compile-time. These \ALLS\ are placed into a compiler-directed memory pool that is
+allocated by the runtime in parallel to FaaS instance startup and before the entry-point function 
+of the workload, $W$, begins. The size of the pool is the sum of all statically 
+analyzable \SIZES\ of the selected \ALLS . As a result, the \ALL\ function calls are 
+replaced by load of a compiler-determined offset of the base address of the pool.
+An \ALL\ can only be selected for this transformation if it is not flagged by 
+the escapes analysis (see the third analysis in \ref{sa}) \textit{and} the compiler 
+can deduce the number of invocations of its parent function. Otherwise, the data
+in memory or the pointer itself may be overwritten accidentally. The latter condition
+can be alleviated with aggressive inlining and function versioning. An example of 
+this transformation is shown in Figure~\ref{fig:pool}. There are a number of 
+benefits to this approach: 1) alleviating function call overheads, 2) enabling 
+prefetching of allocation pointer address, 3) hiding allocation overhead by performing 
+it before $W$ even begins.
+
+\begin{figure}
+    \centering
+    \begin{minipage}{0.45\textwidth}
+        \centering
+        \includegraphics[width=0.77\textwidth]{figs/pool.png} 
+        \caption{Example allocations (top) transformed to fetch from the compiler-directed
+        memory pool (bottom; here, wrapped into the function \texttt{pool\_fetch}). The 
+        basic functionality of the pool is shown to the right. In context of simplified LLVM 
+        IR and \textit{libc} \texttt{malloc}}  
+	    \label{fig:pool}
+    \end{minipage}\hfill
+\end{figure}
 

### 2021-12-09 paper/paper.tex
[paper]: add rest of custom allocator subsection

@@
 
-\subsection{Profiler}
+\subsection{Profiler} \label{prof}
 The profiler has the goal of understanding information about dynamic memory \ALLS\ and 
@@
 \textit{pre-allocated} because all the information needed to allocate is known at
-compile-time. These \ALLS\ are placed into a compiler-directed memory pool that is
+compile-time. These \ALLS\ are placed into a \textit{compiler-directed memory pool} that is
 allocated by the runtime in parallel to FaaS instance startup and before the entry-point function 
@@
 the escapes analysis (see the third analysis in \ref{sa}) \textit{and} the compiler 
-can deduce the number of invocations of its parent function. Otherwise, the data
+can deduce the number of invocations of its parent function or of the allocation
+itself (if it belongs to a loop, for instance). Otherwise, the data
 in memory or the pointer itself may be overwritten accidentally. The latter condition
@@
 prefetching of allocation pointer address, 3) hiding allocation overhead by performing 
-it before $W$ even begins.
+it before $W$ even begins. This can boost performance and simplifies memory management
+for FaaS instances.
+
+Allocations that have statically analyzable size but could not be transformed to
+use the compiler-directed memory pool, can still be transformed to invoke other
+methods of the allocator. We motivate this by a simple example: a node 
+to a linked list is allocated during an irregular while loop. These \ALLS\ most
+likely belong to the same static IR instruction, but are invoked many times with
+the same \SIZE\ during the loop. Consequently, we transform these \ALLS\ to 
+invoke an allocation callback from a \textit{bump allocator} that has a block size equal to the 
+\SIZE . The callback, \texttt{AllocateFromBump(blk\_size)}, returns a pointer from the
+runtime's bump allocator dedicated to a block size of \texttt{blk\_size}. The bump allocator
+for \texttt{blk\_size} is initialized in parallel to FaaS instance startup via 
+a compiler-injected callback, since \texttt{blk\_size} is already known at compile-time.
+All bump allocators are stored by their block sizes in the runtime.
+
+Finally, \ALLS\ that may be invoked many times (as in the example), but have 
+variable-dependent \SIZES\ (i.e. the size is not a constant integer) are transformed to 
+invoke a callback, \texttt{AllocateFromBumpWithRuntimeInit(blk\_size)}, that performs 
+bump allocator initialization during runtime if the allocator for \texttt{blk\_size} is not
+yet initialized. \ALLS\ that fit the criteria for this transformation include ones
+in loops with loop-invariant and variable-dependent \SIZES .
+
+It is important to note that the profiler (see \ref{prof}) can aid these transformations.
+For instance identifying hot functions can expand the number of \ALLS\ that could
+be transformed to invoke \texttt{AllocateFromBumpWithRuntimeInit}, since those static
+\ALLS\ would be executed many times. Moreover, each bump allocator starts with a pool of 
+a default number of of entries to allocate from before expanding to more pools. Using
+profile information, the number of entries could be tuned for each specific bump allocator.
 

### 2021-12-09 paper/paper.tex
[paper]: add implementation

@@
 \section{Implementation}
+The prototype compiler was built using LLVM 13.0.0. Both prototype runtimes (profiler and allocator) 
+were implemented using \texttt{C++}. Profiler-guided compiler instrumentation for the custom 
+allocator was not fully implemented and tested in this prototype due to its engineering effort.
+
+In the allocator runtime in particular, bump allocators are organized in a red-black tree (\texttt{std::map})
+where the key is the block size. Each bump allocator object allocates from an initial pool
+containing (a default) 32 entries (each entry is sized to the block size). If there are no free entries
+in the pool, another pool is allocated. Pools for a bump allocatore are also stored in a 
+red-black tree (\texttt{std::map}), mostly to aid \DALLS .
+
+Initialization for the compiler-directed memory pool and bump allocators with known block size at 
+compile-time are injected by the compiler into a runtime function marked with the attribute 
+\texttt{constructor}. Although initialization is theoretically most performant when done in parallel
+to FaaS startup, this was not implemented in the system prototype.
+
+Finally, all workloads are implemented in \texttt{C} and only use \textit{libc} \texttt{malloc} and 
+\texttt{free}. Although most FaaS workloads are implemented in higher-level languages, the 
+simplicity of the dynamic memory interface makes it very simple for the prototype compiler and 
+runtime to target, as all \ALLS\ and \DALLS\ are obviously explicit. Targeting higher-level 
+languages remains a future work.
 

### 2021-12-09 paper/paper.tex
[paper]: start eval, fix up figure placement, fix content

@@
 \usepackage{amsmath}
+\usepackage{multirow}
 \usepackage{graphicx}
+\usepackage{array}
 \usepackage{titlesec}
@@
 
-\begin{figure}
+\begin{figure}[htp]
     \centerline{\pagefig{figs/sys.pdf}}
@@
         \centering
-        \includegraphics[width=0.77\textwidth]{figs/sizes.png} 
+        \includegraphics[width=0.70\textwidth]{figs/sizes.png} 
         \caption{Example allocations that the static analysis can target in the 
@@
 The allocator primarly takes the static analyses (see \ref{sa})
-as input to replace as many \ALLS\ and \DALLS\ with invocations to the allocator. 
+as input to replace as many \ALLS\ and \DALLS\ with invocations to the allocator. If
+an allocation is not selected as a candidate for any transformations, it is left unchanged.
 First, \ALLS\ with statically analyzable \SIZES\ are prime candidates to be 
@@
 
-\begin{figure}
-    \centering
-    \begin{minipage}{0.45\textwidth}
+\begin{figure} [htp]
         \centering
@@
 	    \label{fig:pool}
-    \end{minipage}\hfill
 \end{figure}
@@
 
-Initialization for the compiler-directed memory pool and bump allocators with known block size at 
-compile-time are injected by the compiler into a runtime function marked with the attribute 
-\texttt{constructor}. Although initialization is theoretically most performant when done in parallel
-to FaaS startup, this was not implemented in the system prototype.
+Initialization for the compiler-directed memory pool and bump allocators (with known block size at 
+compile-time) are injected by the compiler into a runtime function marked with the \texttt{constructor}
+attribute. Although initialization is theoretically most performant when done in parallel
+to FaaS startup, this was not implemented in the prototype.
 
@@
 \section{Evaluation}
+To evaluate the prototype, four kernels (matrix multiplication, compressed-row-storage conversion,
+breadth first traversal, and Huffman coding -- see Figure~\ref{fig:apps}) were used. All testing
+was performed on a 28-Core Intel Xeon CPU E5-2690 v4 machine. All runs were single-threaded.
+
+\begin{figure}
+        \small
+        \centering
+        \begin{tabular}{ | m{5cm} | m{7cm} | } 
+          \hline
+            Matrix Multiplication & Naive algorithm ; [60 x 30] x [30 x 120] \\ 
+          \hline
+            Compressed-Row-Storage (CSR) & Converts a [500 x 400] matrix to CSR form \\ 
+          \hline
+            Breadth First Traversal & Traversal on graph with 1000 vertices and records the traversal ; graph represented with adjancency matrix  \\ 
+          \hline
+            Huffman Coding & Builds a Huffman Tree for a block of "Lorem Ipsum" text \\ 
+          \hline
+        \end{tabular}
+        \caption{Kernels used for evaluation.}
+        \label{fig:apps}
+\end{figure}
+
+We assess the compiler transforms for the custom allocator both statically and dynamically. Figure X
+shows the number of eligible static \ALLS\ in the baseline IR for each kernel and the number of \ALLS\
+that were transformed to invoke the allocator.
+
+Next, we evaluate the performance of the transformed binaries against the baseline binaries (i.e.
+without transformation, compiled straight from source code) for each kernel. Each kernel was 
+run 10 times via \texttt{perf stat} and the cycle counts were recorded. Figure X shows the comparative
+overhead ratio (in terms of cycles).
+
+Finally, we break down the execution times for each major piece of the allocator runtime and
+highlight worst-case observed times (also in cycles). Figure X shows this for each essential
+method of the runtime.
+
+\begin{figure}
+        \small
+        \centering
+        \begin{tabular}{ | m{5cm} | m{9cm} | m{3cm} | } 
+          \hline
+            \texttt{Init} & Initialize the global allocator (that manages bump allocators) and compiler directed pool & 20418 \\ 
+          \hline
+            \texttt{AddAllocator} & Initialize a new bump allocator and add it to the global allocator & 27170 \\ 
+          \hline
+            \texttt{AllocateFromPool} & Fetching a pointer from the compiler-directed memory pool (wrapped as a function) & 52 \\ 
+          \hline
+            \texttt{AllocateFromBump} & Fetching a pointer from a bump allocator & 1372 \\ 
+          \hline
+            \texttt{AllocateFromBumpWithRuntimeInit} & Fetching a pointer from a bump allocator with allocator initialization, if necessary & 1372 \\ 
+          \hline
+        \end{tabular}
+        \caption{Worst-case cycle times for essential methods in allocator runtime.}
+        \label{fig:apps}
+\end{figure}
+
+Note that the profiler, which outputs the \DWS\ and raw data on working set size 
+every allocation/deallocation in the workload, is functional but not shown here.
 

### 2021-12-09 paper/paper.tex
[paper]: add more eval text, figures and tables, add conclusion

@@
 runtime to target, as all \ALLS\ and \DALLS\ are obviously explicit. Targeting higher-level 
-languages remains a future work.
+languages remains a future work. All inputs are placed on the stack as a simulation of the 
+JIT/binary translation-based solution for embedding inputs into the binary before a FaaS 
+instance executes.
 
@@
 
-We assess the compiler transforms for the custom allocator both statically and dynamically. Figure X
-shows the number of eligible static \ALLS\ in the baseline IR for each kernel and the number of \ALLS\
-that were transformed to invoke the allocator.
+We assess the compiler transforms for the custom allocator both statically and dynamically. Figure~\ref{fig:static}
+shows the number of eligible static \ALLS\ (\texttt{malloc}s column) in the baseline IR for each kernel and 
+the number of \ALLS\ that were transformed to invoke the allocator ("Transformed" column). For matrix 
+multiplication, the compiler aggressively unrolled the entire allocation loop for the result matrix
+and inlined the entire \texttt{multiply} function. Consequently, all \SIZES\ were statically analyzable
+and all \ALLS\ were transformed to fetch from the compiler-directed pool. For breadth first traversal and Huffman 
+coding, almost all \ALLS\ were transformed to allocate from the bump allocator because most \ALLS\
+were found in loops but had statically analyzable \SIZES . The \ALLS\ not transformed in
+the Huffman coding kernel were all in functions that were inlined. The inlined static \ALLS\ were
+all transformed.
+
+\begin{figure}[htp]
+    \centering
+    \begin{minipage}{0.45\textwidth}
+        \small
+        \centering
+        \begin{tabular}{ | m{5cm} | m{1cm} | m{2cm} | } 
+          \hline
+            Kernel & \texttt{malloc}s & Transformed \\
+          \hline\hline
+            Matrix Multiplication & 60 & 60 \\ 
+          \hline
+            Compressed-Row-Storage (CSR) & 3 & 0 \\ 
+          \hline
+            Breadth First Traversal & 7 & 7  \\ 
+          \hline
+            Huffman Coding & 17 & 10 \\ 
+          \hline
+        \end{tabular}
+        \caption{Kernels used for evaluation.}
+        \label{fig:static}
+    \end{minipage}\hfill
+    \begin{minipage}{0.45\textwidth}
+        \centering
+        \includegraphics[width=0.88\textwidth]{figs/graph.png} 
+        \caption{Overhead ratio for transformed kernels (with allocator runtime)
+        vs. baseline kernels (no transform). Ratio compares cycle counts.}  
+	    \label{fig:perf}
+    \end{minipage}\hfill
+\end{figure}
 
@@
 without transformation, compiled straight from source code) for each kernel. Each kernel was 
-run 10 times via \texttt{perf stat} and the cycle counts were recorded. Figure X shows the comparative
-overhead ratio (in terms of cycles).
+run 10 times via \texttt{perf stat} and the cycle counts were recorded. Figure~\ref{perf} shows the 
+comparative overhead ratio (in terms of cycles). Here, all kernels have minimal overhead; although
+matrix multiplication, in particular, should be faster. The lack of speedups could be attributed
+to a very low-latency \texttt{malloc} coupled with extra initialization and runtime overhead. This
+is explored in Figure~\ref{cycles}. We break down the execution times for each major piece of the 
+allocator runtime and highlight worst-case observed times (also in cycles). These overheads are
+fairly substantial for the work that each method is performing. Some of the overhead can be due
+to extra function calls because wrapper methods need to be exposed to the compiler for streamlined
+instrumentation. Additionally, the data structures (\texttt{std::map}) used in the runtime and 
+initialization costs may drive up overheads. Many of these overheads could be optimized by 
+careful engineering.
 
-Finally, we break down the execution times for each major piece of the allocator runtime and
-highlight worst-case observed times (also in cycles). Figure X shows this for each essential
-method of the runtime.
-
-\begin{figure}
+\begin{figure} [htp]
         \small
         \centering
-        \begin{tabular}{ | m{5cm} | m{9cm} | m{3cm} | } 
+        \begin{tabular}{ | m{5cm} | m{9cm} | m{1cm} | } 
           \hline
@@
         \caption{Worst-case cycle times for essential methods in allocator runtime.}
-        \label{fig:apps}
+        \label{fig:cycles}
 \end{figure}
@@
 
-\section{Future Work}
-
 \section{Conclusion}
+This paper shows a \crcd\ that analyzes and optimizes for dynamic memory usage in 
+small FaaS workloads. The system provides a profiler, static allocation tracking and 
+analysis, and a custom memory allocator controlled by both the compiler and runtime
+tooptimize for dynamic memory. Across a small set of kernels, the compiler is 
+successful at analyzing and transforming allocation candidates to use the custom
+allocator, and the system achieves comparable performance compared to the baseline 
+applications.
 

### 2021-12-09 paper/paper.tex
[paper]: make edits

@@
 
-\section{Introduction/Background}
-
 \section{Design}
@@
 custom allocator replaces standard dynamic \ALLS\ and \DALLS\ in \texttt{app.c}, and it is
-the \textit{crux} of optimization for dynamic memory usage in FaaS workloads. The
+the \textit{crux} of optimization for dynamic memory usage in FaaS workloads in this system. The
 outputs from this compilation pipeline include an instrumented application (\texttt{app.exe})
@@
 The profiler has the goal of understanding information about dynamic memory \ALLS\ and 
-\DALLS\ during execution (for instance, the number of \ALLS\, \DWS\, common \ALL 
+\DALLS\ during execution (for instance, the number of \ALLS\, \DWS\, common \ALL\ 
 sizes, etc.) as well as typical information about critical paths or the "hotness" of functions and
-loops at runtime that is often use to aid optimization [?]. Using this data, the profiler can inform the compiler during 
+loops at runtime that is often use to aid optimization. Using this data, the profiler can inform the compiler during 
 instrumentation in the second pass over the application (see \ref{ca}). It can also 
@@
 
-The profiler tracks all \ALLS and \DALLS of some workload, $W$, at runtime. This is 
+The profiler tracks all \ALLS\ and \DALLS\ of some workload, $W$, at runtime. This is 
 accomplished by using the compiler to instrument all explicit dynamic memory \ALL\ and \DALL\ function 
@@
 The system also supports static analyses of \ALLS\ and \DALLS\ in the IR for workload $W$. 
-In particular, the compiler analyzes 1) \ALL\ sizes for all \ALLS\, 2) loops to unroll 
+In particular, the compiler analyzes 1) \ALL\ sizes for all \ALLS\ , 2) loops to unroll 
 and functions to inline so that 1) is more feasible to analyze, and 3) uses of \ALL\ and 
@@
 inline the entry-point function of $W$. This is especially common in small 
-FaaS workloads where \ALLS\ towards the beginning of the workload havew \SIZES\
+FaaS workloads where \ALLS\ towards the beginning of the workload have \SIZES\
 dependent on the inputs. Obviously, inlining the entry-point function is not 
@@
 to possibly embed the inputs into the binary itself. This process can occur
-right before the $W$ is invoked in the FaaS instance. A subset of this system's compiler
+right before $W$ is invoked in the FaaS instance. A subset of this system's compiler
 can analyze and transform $W$ as necessary. This solution is theoretical and 
@@
 determine an appropriate unroll factor (unrolling too much without knowledge of
-the \SIZE\ may be sub-optimal). An example of this is shown in Line 8 of Figure ~\ref{fig:sizes}
+the \SIZE\ may be sub-optimal). An example of this is shown in Line 8 of Figure ~\ref{fig:sizes}.
 
@@
 are recorded. All \ALLS\ and \DALLS\ that cannot be statically analyzed using
-memory alias analysis (i.e "must alias" conditions) or def-use chains are flagged.
+memory alias analysis (i.e "must alias" conditions) or def-use chains are conservatively flagged.
 
@@
 In particular, the allocator aims to reduce raw \ALL\ and \DALL\ time, consolidate 
-\ALLS\ , and utilize compiler analysis to optimize dynamic memory overheads.
+\ALLS\, and utilize compiler analysis to optimize dynamic memory overheads.
 
@@
 analyzable \SIZES\ of the selected \ALLS . As a result, the \ALL\ function calls are 
-replaced by load of a compiler-determined offset of the base address of the pool.
+replaced by a load of a compiler-determined offset of the pool's base address.
 An \ALL\ can only be selected for this transformation if it is not flagged by 
@@
 benefits to this approach: 1) alleviating function call overheads, 2) enabling 
-prefetching of allocation pointer address, 3) hiding allocation overhead by performing 
+prefetching of allocation pointer address, and 3) hiding allocation overhead by performing 
 it before $W$ even begins. This can boost performance and simplifies memory management
@@
 the same \SIZE\ during the loop. Consequently, we transform these \ALLS\ to 
-invoke an allocation callback from a \textit{bump allocator} that has a block size equal to the 
+invoke an allocation callback from a fast \textit{bump allocator} that has a block size equal to the 
 \SIZE . The callback, \texttt{AllocateFromBump(blk\_size)}, returns a pointer from the
@@
 bump allocator initialization during runtime if the allocator for \texttt{blk\_size} is not
-yet initialized. \ALLS\ that fit the criteria for this transformation include ones
+yet initialized. Allocations that fit the criteria for this transformation include ones
 in loops with loop-invariant and variable-dependent \SIZES .
@@
 It is important to note that the profiler (see \ref{prof}) can aid these transformations.
-For instance identifying hot functions can expand the number of \ALLS\ that could
+For instance, identifying hot functions can expand the number of \ALLS\ that could
 be transformed to invoke \texttt{AllocateFromBumpWithRuntimeInit}, since those static
@@
         basic functionality of the pool is shown to the right. In context of simplified LLVM 
-        IR and \textit{libc} \texttt{malloc}}  
+        IR and \textit{libc} \texttt{malloc}.}  
 	    \label{fig:pool}
@@
 instance executes.
+
+The prototype can be found at \href{https://github.com/sgh185/849}{https://github.com/sgh185/849}.
 
@@
         \end{tabular}
-        \caption{Kernels used for evaluation.}
+        \caption{Number of static allocations transformed.}
         \label{fig:static}
@@
 without transformation, compiled straight from source code) for each kernel. Each kernel was 
-run 10 times via \texttt{perf stat} and the cycle counts were recorded. Figure~\ref{perf} shows the 
+run 10 times via \texttt{perf stat} and the cycle counts were recorded. Figure~\ref{fig:perf} shows the 
 comparative overhead ratio (in terms of cycles). Here, all kernels have minimal overhead; although
@@
 to a very low-latency \texttt{malloc} coupled with extra initialization and runtime overhead. This
-is explored in Figure~\ref{cycles}. We break down the execution times for each major piece of the 
+is explored in Figure~\ref{fig:cycles}. We break down the execution times for each major piece of the 
 allocator runtime and highlight worst-case observed times (also in cycles). These overheads are
